\chapter{Data Required}
\label{ch:data-required}

• what data were needed for the project and where it was obtained from; \\
• ethical use of data, including use of human data & human participants: \\
1. ethical use of data \\
! explicitly specify whether you used \\
! Synthetic data, \\
or \\
! Real Non Human data \\
! explicitly confirm an ethical source of the data, \\
! confirm the University or a relevant Professional Body \\
Ethical approval has been obtained for the use of the
data in your project. \\
! where applicable, include into appendix the University \\
Ethical approval obtained by your 1st supervisor for the
project on your behalf. \\
or \\
! Real Human data: \\
! explicitly confirm an ethical source of the data \\
! explicitly confirm that the University Policy on ethical \\
use of human data has been followed: here is the flow
chart for the University Ethical approval. \\
! explicitly confirm that the University or a relevant
Professional Body Ethical approval has been obtained
for the use of the data in your project. \\
! where applicable, include into appendix the University
Ethical approval obtained by your 1st supervisor for the
project on your behalf \\
! be aware that only the following types of data do not require
Research ethics approval: \\
1. information freely available in the public domain; \\
2. anonymised records and data sets that exist in the public
domain \\
2. ethical use of human participants (other then project 3rd party evaluation) \\
! explicitly state if human participants were involved in the project; \\
! if human participants were involved in the project, \\
! explicitly confirm that the University ethical procedure has been
followed: here is the flow chart for the University Ethical
approval; \\
! explicitly confirm that the University or a relevant Professional
Body Ethical approval has been obtained for the use of human
participants in your project; \\
! submit to your project 1st supervisor the hard copy originals of
the consent forms signed by the human participants on your
project, for the subsequent 1 year storage at the Departmental
Student Office. \\
! for completeness of your project report, you might consider to
include into the appendices: \\
! blank documents you have developed for your project
human participants, namely: your project information
sheet, the questionnaire(s), and your project human
participant consent form, \\
! if applicable and necessary, then the
fully anonymised copies of signed consent forms. \\

\section{Synthetic Data}

% Ignore this for now
For the input data, several entries from the Cambridge Structural Database
were used. For testing, I’ve made use of Mercury, a visualisation tool for
crystal structures that works in unison with the CSD. To use Mercury and
access the CSD as a resource, a license is required. For the purpose of this
project, I have been granted a license by Dr Vitaliy Kurlin from the University
of Liverpool. As integration with the CSD has not been achieved in time, the
user does not require a license for the CSD to utilise the project, with the
exception if they choose to use it as a means of extracting crystalline material
data. In that event, a license will still be required. \\
A lot of caution has been utilised when selecting 3rd party libraries, and
all used have been check and proven to be open-source. No further user
attention is required as the program comes packed with all the required
library files under a Dynamic-Linked Library (DLL) format. \\