\chapter{Evaluation}
\label{ch:evaluation}

This gives an evaluation of the project, including \\
• A description of how the project is evaluated, including \\
o What criteria are used to evaluate whether the system is suceesful? \\
o How these criteria are assessed? \\
o Who is involved in the evaluation? \\
• Your critical evaluation of your project results/outcomes. \\
• Your critical evaluation of the strengths and weaknesses of your project as carried out. \\
• Where appropriate, 3rd party evaluation of the software/computer
system/application developed on the project, and/or customer feedback, be
obtained in strict accordance with ethical use of the project 3rd party evaluation human participants or explicitly state if human participants were involved for the project 3rd party evaluation; \\
o if human participants were involved for the project 3rd party evaluation, \\
! explicitly confirm that the CS Department ethical procedure for MEng
projects 3rd party evaluation has been followed. \\
! submit to your project 1st supervisor the hard copy originals of the
consent forms signed by 3rd party evaluators on your project, for the
subsequent 1 year storage at the Departmental Student Office. \\
! for completeness of your project report, you might consider to include
into the appendices: \\
! blank documents you have developed for your project 3rd
party evaluation, namely: blank copies of your project
information sheet, consent form, and questionnaire(s), \\
! if applicable and necessary, then the fully anonymised copies
of questionnaire(s), and/or fully anonymised copies of the
signed consent forms. \\

%%%%%%%%%%%%%%%%%%%%%
The system will mainly be evaluated by personnel in the MIF as they have collectively commissioned the system to be build, participation in this evaluation will be optional and informal aside from a selected group from the project group itself. With this select group we will collect more in-depth data in the form of questionnaires. \\

\section{Criteria of ass-essment ;^)}
\section{Test case: LiAlSO}
\section{Test case: insert formula I}
\section{Test case: insert formula II}
\section{Test case: insert formula III}
\section{Strengths and Weaknesses}